---
title: "Modern Applied Statistics exercises from ISLR"
author: "Yamuna Dhungana"
output: 
    pdf_document:
      latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,warning=F,message=F)
```


### Use set.seed(20218) in each exercise to make results reproducible.

### Use 1,000 bootstrap samples where bootstrap is required.

1. Question 5.4.2 pg 197. *Justify your answers and spend some time thinking about the implications of these experiments.*
We will now derive the probability that a given observation is part
of a bootstrap sample. Suppose that we obtain a bootstrap sample
from a set of n observations.

a. What is the probability that the first bootstrap observation is
not the jth observation from the original sample? Justify your
answer.

   *In a bootstrap, the sample is taken with replacement from the total sample (n) therefore, probability
   of being jth observation is 1/n hence probability of not being jth observation is 1-1/n*


b. What is the probability that the second bootstrap observation
is not the jth observation from the original sample?

   *For all the samples, data is taken randomly with replacement therefore, the probability of observation 
   that is not jth observation will not be different from the first one. The probability of not being jth
   observation will be $1-1/n$*


c. Argue that the probability that the jth observation is not in the
bootstrap sample is $(1 − 1/n)^n$.

   *We know that the n is the total observation and the probability of not being the jth observation is 
   $1-1/n$. The oribability that the jth observation is not in the bootstrap sammple is $(1 − 1/n)^n$. 
   Here, we multiplied the n because the sample is taken without replacement and is are an independent 
   sample. 
   $(1−1/n)⋯(1−1/n)=(1−1/n)^n$*


d. When n = 5, what is the probability that the jth observation is
in the bootstrap sample?

   *The probability that the probability that the Jth observation is in the bootstrap sample is $1-(1-1/n)^n$
   
   Now,
   calculating with n=5
      p(j belong in bootstrap sample) $= 1-(1-1/n)^n$
                                      $= 1-(1-1/5)^5$
                                      $= 0.67232$

   Hence, The probabilty that the Jth observation is in boot strap sample is 0.67232*
 

e. When n = 100, what is the probability that the jth observation
is in the bootstrap sample?

   *The probability that the probability that the Jth observation is in the bootstrap sample is $1-(1-1/n)^n$
   
   Now,
   calculating with n=100
      p(j belong in bootstrap sample) $= 1-(1-1/n)^n$
                                      $= 1-(1-1/100)^{100}$
                                      $= 0.6339677$

   Hence, The probabilty that the Jth observation is in boot strap sample is 0.634*
 
      

f. When n = 10,000, what is the probability that the jth observation
is in the bootstrap sample?

   
   *The probability that the probability that the Jth observation is in the bootstrap sample is
   $1-(1-1/n)^n$
   
   calculating with n=10000
   
   p(j belong in bootstrap sample) $= 1-(1-1/n)^n$
                                   $= 1-(1-1/10000)^{10000}$
                                   $ = 0.632139$

   Hence, The probabilty that the Jth observation is in boot strap sample is 0.632*



g. Create a plot that displays, for each integer value of n from 1
to 100,000, the probability that the jth observation is in the
bootstrap sample. Comment on what you observe.

```{r}
library(ggplot2)
n <- c(1:100000)
p <- 1-(1-1/n)^n
mydata <- data.frame(n, p)
ggplot(mydata, aes(x=n, y=p)) + 
  geom_line(size=0.5, col= "blue") + 
  xlim(c(0, 100)) + ylim(c(0.5, 1))+
  labs(title="Probability that the number belong in jth observation", y="Probability", x="Number")

```
   
   *Above is the graphical representation of each integer value of n from 1 to 100,000 and the 
   probability that the jth observation is in the bootstrap sample. Here the line continues to 
   100000 number from number 12.5. Here, I have limited the x-axis because I wanted to see the 
   curve, which we see in the graphs before 12.5. Without zooming in the graph, it appears a
   straight line starting from 6.3 on the y-axis. The highest probability of the number is one
   and, the lowest is near 0.63.*


h. We will now investigate numerically the probability that a bootstrap
sample of size n = 100 contains the jth observation. Here
j = 4. We repeatedly create bootstrap samples, and each time
we record whether or not the fourth observation is contained in
the bootstrap sample.

```{r}
set.seed(20218)
store=rep (NA , 10000)
for (i in 1:10000) {
store[i]=sum(sample (1:100 , rep =TRUE)==4) >0
}
mean(store)
```
   
   *Here, We are trying to obtain a bootstrap sample of size n = 100 that contains the jth 
   observation. Here j = 4. We repeatedly create bootstrap samples, and each time we record 
   whether or not the fourth observation is in the bootstrap sample. We have created a bootstrap
   sample of 100 samples repeating with a replacement for 10000 times using for loop. The statistic
   (mean) is performed of the sample that has the fourth observation. The result is reported. Here,
   The mean of the data that satisfied the condition is 0.6316, which means 63.16% of the samples 
   had the fourth observation. Here I added set.seed in my answer for my convenience.*




2. Question 5.4.9 pg 201. *For this question, do not use the* **boot** *library or similar functions. You are expected to code it up in base R with formal annotated code.*

We will now consider the Boston housing data set, from the MASS
library.
a. Based on this data set, provide an estimate for the population
mean of medv. Call this estimate ˆμ.

```{r}
library(MASS)
data(Boston)
# head(Boston)

muhat <- mean(Boston$medv)
paste0("The standard error of mu is ", round(muhat, digits = 3))
```

   *The question is straightforward and is the estimation of the population mean is the sample 
   mean. The estimated population mean of medv was found to be 22.53281.*
   

b. Provide an estimate of the standard error of ˆμ. Interpret this
result.
Hint: We can compute the standard error of the sample mean by
dividing the sample standard deviation by the square root of the
number of observations.

```{r}
std_error <- (sd(Boston$medv))/sqrt(dim(Boston)[1])
paste0("The standard error of mu is ", round(std_error, digits = 6))

```
   
   * By using the hint we found out that the standard error of the mu is 0.408861.*
   

c. Now estimate the standard error of ˆμ using the bootstrap. How
does this compare to your answer from (b)?

```{r}
set.seed(20218)
n <- dim(Boston)[1]
B <- 1000
boot.samples <- matrix(sample(Boston$medv, size = B * n, replace = TRUE),
                      B, n)
# dim(boot.samples)
boot.statistics = apply(boot.samples, 1, mean)

require(ggplot2)
ggplot(data.frame(medv = boot.statistics),aes(x=medv)) +
  geom_histogram(binwidth=0.25,aes(y=..density..), col="grey") +
  geom_density(color="red")

medv.se = sd(boot.statistics)
paste0("The standard error of medv with bootstrap is ", round(medv.se, digits = 6))

```
   
  *Here, I have set the seed with 20218 and the bootstrap number of 1000 as mentioned by the 
  question. I have plotted the density plot of the mean of these data. The standard error in
  bootstrap is calculated by the standard deviation. The standard error of medv with bootstrap
  is 0.417868. The error obtained by bootstrap is a bit large (about 1 %) than the error obtained
  without bootstrap.*
  

d. Based on your bootstrap estimate from (c), provide a 95% confidence
interval for the mean of medv. Compare it to the results
obtained using t.test(Boston$medv).
Hint: You can approximate a 95% confidence interval using the
formula [ˆμ − 2SE(ˆμ), ˆμ + 2SE(ˆμ)].
```{r}
t.test(Boston$medv)

confidence95 <- (c(muhat - 2*medv.se, muhat + 2*medv.se))
confidence95


```

   *By using the hint given, I have calculated the confidence interval of 95 percentile. 
   The result was obtained by the t-test and the given hint. I have obtained the same 
   confidence with one decimal point of intervals 21.7 and 23.3.*


e. Based on this data set, provide an estimate, ˆμmed, for the median
value of medv in the population.

```{r}
medhat <- median(Boston$medv)
paste0("The standard error of medv with bootstrap is ", round(medhat, digits = 6))

```


f. We now would like to estimate the standard error of ˆμmed. Unfortunately,
there is no simple formula for computing the standard
error of the median. Instead, estimate the standard error of the
median using the bootstrap. Comment on your findings.

```{r}
set.seed(20218)
n <- dim(Boston)[1]
B <- 1000
boot.samples <- matrix(sample(Boston$medv, size = B * n, replace = TRUE),
                      B, n)
dim(boot.samples)
boot.statistics.median = apply(boot.samples, 1, median)

require(ggplot2)
ggplot(data.frame(medv = boot.statistics.median),aes(x=medv)) +
  geom_histogram(binwidth=0.25,col="grey") +
  geom_density(color="red")

medv.se.med = sd(boot.statistics.median)
paste0("The standard error of medv with bootstrap is ", round(medv.se.med, digits = 6))


```

   *Like earlier, I have repeated the same process as before the only difference from before
   is median, is preformed here. Here, I have set the seed with 20218 and the bootstrap number
   of 1000 as mentioned by the question. I have plotted the bar chat of the median of these 
   data. The standard error in bootstrap is calculated by the standard deviation. The standard
   error of medv with bootstrap is 0.384951. The error obtained by bootstrap is 0.384951 which 
   is small.*
   
   
g. Based on this data set, provide an estimate for the tenth percentile
of medv in Boston suburbs. Call this quantity ˆμ0.1. (You
can use the quantile() function.

```{r}
muhat.1=quantile(Boston$medv, probs=0.1)
paste0("The tenth percentile of medv is ", muhat.1)

```


h. Use the bootstrap to estimate the standard error of ˆμ0.1. Comment
on your findings.

```{r}

n <- dim(Boston)[1]
B <- 1000
set.seed(20218)
boot.samples <- matrix(sample(Boston$medv, size = B * n, replace = TRUE),
                      B, n)

boot.statistics.mu = apply(boot.samples,1, mean)

est <- boot.statistics.mu
# perc <- c(quantile(boot.samples, prob=0.1))

muhat0.1.se <- (2*est-quantile(boot.samples, prob=0.1))

# standard error for the mu 0.1
paste0("The tenth percentile of medv is ", (round(sd(muhat0.1.se), digits = 5)))

```
   
   *Like earlier, I have repeated the same process as before.Here, I have set the seed with 
   20218 and the bootstrap number of 1000 as mentioned by the question.I have performed with
   the mean. The confidence interval is performed and The standard error in bootstrap is 
   calculated by the standard deviation. The standard error of medv with bootstrap is 0.384951.
   The error obtained by bootstrap is 0.835736 which is still small.*
   

```{r}
# https://www.youtube.com/watch?v=O_Fj4q8lgmc&list=PLqzoL9-eJTNDp_bWyWBdw2ioA43B3dBrl&index=1
# https://stats.idre.ucla.edu/r/library/r-library-introduction-to-bootstrapping/
# https://stackoverflow.com/questions/48665924/r-bootstrap-percentile-confidence-interval
# http://pages.stat.wisc.edu/~larget/stat302/chap3.pdf
# https://www.youtube.com/watch?v=Om5TMGj9td4&list=PLqzoL9-eJTNDp_bWyWBdw2ioA43B3dBrl&index=5
# https://www.statisticshowto.com/probability-and-statistics/statistics-definitions/what-is-the-standard-error-of-a-sample/
# ISLR Book 

```

